The role of priors, competence and relevance for the interpretation of disjunction \thanks{We thank \ldots}

### 2022-05-11 analysis/main/02_xor-some_prereg_plots_final.Rmd
plot w means first data

Read data created in preprocessing script:
```{r, warning=FALSE, message=FALSE, results='hide'}
d_critical_zScored_wide <- read_csv("./../../data/main/results_prereg_tidy_final_zScored_wide.csv")
d_critical_zScored <- read_csv("./../../data/main/results_prereg_tidy_final_zScored_long.csv")
d_test <- read_csv("./../../data/main/results_prereg_raw_final_test.csv")
d_critical_zScored_wide_xor_priors_order <- read_csv("./../../data/main/results_prereg_tidy_final_zScored_wide_xor_priors-order.csv")
d_critical_zScored_wide_xor_priors_type <- read_csv("./../../data/main/results_prereg_tidy_final_zScored_wide_xor_priors-type.csv")
```

@@

>>>>

Read data created in preprocessing script:
```{r, warning=FALSE, message=FALSE, results='hide'}
d_critical_zScored_wide <- read_csv("./../../data/main/results_prereg_tidy_final_zScored_wide.csv")
d_critical_zScored <- read_csv("./../../data/main/results_prereg_tidy_final_zScored_long.csv")
d_test <- read_csv("./../../data/main/results_prereg_raw_final_test.csv")
d_critical_zScored_wide_xor_priors_order <- read_csv("./../../data/main/results_prereg_tidy_final_zScored_wide_xor_priors-order.csv")
d_critical_zScored_wide_xor_priors_type <- read_csv("./../../data/main/results_prereg_tidy_final_zScored_wide_xor_priors-type.csv")

d_critical_zScored_meanFirst_noDupl_zScored_wide <- 
  read_csv("./../../data/main/results_prereg_tidy_noDuplicates_zScored_meansFirst_wide.csv")

@@

gridExtra::grid.arrange(p.rel.z, p.comp.z, p.pri.z, ncol = 3) 
```
Replicate plot above on data where means were computed over the prior ratings for or before z-scoring:
```{r}
d_critical_zScored_meanFirst_noDupl_zScored_wide %>%
  ggplot(., aes(x = relevance, y = target)) +
  geom_point(size = 2, alpha = 0.7) +
  ylab("Inference strength ratings") +
  geom_smooth(method = "lm") +
  facet_wrap(~main_type, ncol = 1) -> p.rel.z

d_critical_zScored_meanFirst_noDupl_zScored_wide %>%
  ggplot(., aes(x = competence, y = target)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm") +
  ylab("") +
  facet_wrap(~main_type, ncol = 1) -> p.comp.z

d_critical_zScored_meanFirst_noDupl_zScored_wide %>%
  ggplot(., aes(x = prior, y = target)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm") +
  ylab("") +
  facet_wrap(~main_type, ncol = 1) -> p.pri.z

### 2022-05-11 writing/paper/paper_draft.tex
plot w means first data

% Optional short author inside square brackets, for the running headers.
% If no short author is given, no authors print in the headers.
\author[]{% As many authors as you like, each separated by \AND.
  \spauthor{Polina Tsvilodub \\ \institute{Institute}} \AND
  \spauthor{Bob van Tiel \\ \institute{Institute}} \AND
  \spauthor{Michael Franke \\ \institute{Department of Linguistics \\ University of Tübigen}}}

@@

Participants’ factor ratings by-story agreed well with the designed classification of the stories (Fig. \ref{main-raw-ratings}, red vs. blue color on $x$-axis). That is, the ratings for the relevance, comptence and prior statements were not distributed uniformly across the stories, but align with the prior categorization of the story for the respective factor (Fig. \ref{main-raw-ratings}). This provides validation for our experimental manipulation of the explanatory factors.

@@

We focus on six slope coefficients of the regression model, for the effects of relevance, competence and prior, respectively, once for ``some'' and once for ``or''. We check whether the posterior estimate of each effect is in either one of three intervals: (1) negative effect ($\le$ -0.05), (2) no effect (between -0.05 and 0.05), or (3) positive effect ($\ge$ 0.05). We set the threshold for considering an effect as positive or negative to ±0.05 because we consider 0.05 to be the Region of Practical Equivalence (ROPE) for the effect sizes we expect \citep{kruschke2014doing}.
We interpret the data as providing evidence in favor of an effect (positive, negative, no effect) if the posterior probability of the effect being true is >= 0.95 (i.e., 95\% of posterior samples are in the corresponding interval). The probabilities of the respective coefficients lying in a particular interval are reported below (i.e., for instance, if $P=0.95$ is reported, it means that 95\% of the posterior samples of the given coefficient are in the respective interval). In particular, we speak of evidence in favor of the scalar implicature account of the exclusive disjunction interpretation if the posterior probability of the prior effect being negative, the relevance and competence effects being positive is $\ge$ 0.95.

\begin{figure}[h]
	\begin{center}
		\includegraphics[width=1\linewidth]{images/posterior-effects-main.pdf}
	\end{center}
	\vspace{-0.3cm}
	\caption{Main analysis results: Distributions of the posterior samples of each factor of interest for the trigger ``some'' and ``or''. The colors and the vertical dashed lines indicate the effect intervalls (from left to right: negative, no, positive effect). Points indicate posterior means, thick lines indicate 66\% credible intervals, thin lines indicate 95\% credible intervals.}
	\label{posteriors-main}
\end{figure}

\begin{table}
  \centering
   \begin{tabular}{llccc}
           &            & negative & no effect & positive \\ \midrule
    {some} & relevance  & ...      & ...       & ...      \\
           & competence & 0        & 0         & 1        \\
           & prior      & 0        & 0.001     & 0.999    \\
    {or}   & relevance  & ...      & ...       & ...      \\
           & competence & ...      & ...       & ...      \\
           & prior      & ...      & ...       & ...      \\
   \end{tabular}
  \caption{Results of confirmatory analysis. \mf{Fill me, please.}}
  \label{tab:results-confirmatory}
\end{table}

Results from this analysis are summarized in Table
Consistent with predictions, for the trigger ``some'', we find a clear negative prior probabilty effect of the stronger alternative ``all'' being true, as inidcated by the probability of the negative effect of prior being $P = 0.999$ (Fig.\ref{posteriors-main}, \texttt{Prior~(some)} line, orange color). Similarly, we find a clear positive effect for speaker competence ($P =  1$, Fig.\ref{posteriors-main}, \texttt{Competence~(some)} line, green color). However, we do not find any clear effects of relevance (Fig.\ref{posteriors-main}, \texttt{Relevance~(some)} line, split colors). If anything, the data supports the hypothesis that the effect is not negative.

@@

\begin{figure}[h]
	\begin{center}
		\includegraphics[width=1\linewidth]{images/posterior-effects-categorical.pdf}
	\end{center}
	\vspace{-0.3cm}
	\caption{Exploratory analysis with categorical predictors: Distributions of the posterior samples of each factor of interest for the trigger ``some'' and ``or''. The colors and the vertical dashed lines indicate the effect intervalls (from left to right: negative, no, positive effect). Points indicate posterior means, thick lines indicate 66\% credible intervals, thin lines indicate 95\% credible intervals. \mf{I'm not sure if we should blindly reuse a threshold of 0.05 here without further justification. Something to think about.}}
	\label{posteriors-cat}
\end{figure}

@@

\begin{figure}[h]
	\begin{center}
		\includegraphics[width=1\linewidth]{images/posterior-effects-means.pdf}
	\end{center}
	\vspace{-0.3cm}
	\caption{Exploratory analysis with predictor ratings averaged by-vignette: Distributions of the posterior samples of each factor of interest for the trigger ``some'' and ``or''. The colors and the vertical dashed lines indicate the effect intervalls (from left to right: negative, no, positive effect). Points indicate posterior means, thick lines indicate 66\% credible intervals, thin lines indicate 95\% credible intervals. \mf{I'm not sure if we should blindly reuse a threshold of 0.05 here without further justification. Something to think about.}}
	\label{posteriors-mean}
\end{figure}

@@

>>>>

% Optional short author inside square brackets, for the running headers.
% If no short author is given, no authors print in the headers.
\author[]{% As many authors as you like, each separated by \AND.
  \spauthor{Polina Tsvilodub \\ \institute{Institute}} \AND
  \spauthor{Bob van Tiel \\ \institute{Institute}} \AND
  \spauthor{Michael Franke \\ \institute{Department of Linguistics \\ University of Tübingen}}}

@@

\begin{figure}[h]
	\begin{center}
		\includegraphics[width=1\linewidth]{images/idealized_data_N200_b015.pdf}
	\end{center}
	\vspace{-0.3cm}
	\caption{Simulated ratings for relevance, competence and prior statements (x-axis) in relation to ratings for the strength of pragmatic enrichments (y-axis). Ratings for stories categorized as low (red) w.r.t. a given factor are on average lower (x-axis) than for those categorized as high (blue).}
	\label{main-sim-ratings}
\end{figure}

Participants’ factor ratings by-story agreed well with the designed classification of the stories (Fig. \ref{main-raw-ratings}, red vs. blue color on $x$-axis). That is, the ratings for the relevance, comptence and prior statements were not distributed uniformly across the stories, but align with the prior categorization of the story for the respective factor (Fig. \ref{main-raw-ratings}). This provides validation for our experimental manipulation of the explanatory factors, as can be seen when comparing Fig. \ref{main-raw-ratings} with Fig. \ref{main-sim-ratings} which depicts idealized data (N=200) simulating perfectly borne out hypotheses. \pt{Like this?}

@@

We focus on six slope coefficients of the regression model, for the effects of relevance, competence and prior, respectively, once for ``some'' and once for ``or''. We check whether the posterior estimate of each effect is in either one of three intervals: (1) negative effect ($\le$ -0.05), (2) no effect (between -0.05 and 0.05), or (3) positive effect ($\ge$ 0.05). We set the threshold for considering an effect as positive or negative to ±0.05 because we consider 0.05 to be the Region of Practical Equivalence (ROPE) for the effect sizes we expect \citep{kruschke2014doing}.
We interpret the data as providing evidence in favor of an effect (positive, negative, no effect) if the posterior probability of the effect being true is >= 0.95 (i.e., 95\% of posterior samples are in the corresponding interval). The probabilities of the respective coefficients lying in a particular interval are reported below (i.e., for instance, if $P=0.95$ is reported, it means that 95\% of the posterior samples of the given coefficient are in the respective interval). In particular, we speak of evidence in favor of the scalar implicature account of the exclusive disjunction interpretation if the posterior probability of the prior effect being negative, the relevance and competence effects being positive is $\ge$ 0.95. Example simulated posterior distributions over effect size samples which would confirm all our hypotheses can be seen in Fig. \ref{posteriors-sim}.

\begin{figure}[h]
	\begin{center}
		\includegraphics[width=1\linewidth]{images/posterior-effects-idealized-sim.pdf}
	\end{center}
	\vspace{-0.3cm}
	\caption{ \pt{How's this?} Simulated main analysis results which would confirm all hypotheses: Distributions of the posterior samples of each factor of interest for the trigger ``some'' and ``or''. The effect size was set to 0.25 and matches Fig. \ref{main-sim-ratings} for visualization purposes. The colors and the vertical dashed lines indicate the effect intervalls (from left to right: negative, no, positive effect). Points indicate posterior means, thick lines indicate 66\% credible intervals, thin lines indicate 95\% credible intervals.}
	\label{posteriors-sim}
\end{figure}

\begin{figure}[h]
	\begin{center}
		\includegraphics[width=1\linewidth]{images/posterior-effects-noDupl.pdf}
	\end{center}
	\vspace{-0.3cm}
	\caption{Main analysis results: Distributions of the posterior samples of each factor of interest for the trigger ``some'' and ``or''. The colors and the vertical dashed lines indicate the effect intervalls (from left to right: negative, no, positive effect). Points indicate posterior means, thick lines indicate 66\% credible intervals, thin lines indicate 95\% credible intervals.}
	\label{posteriors-main}
\end{figure}

\begin{table}
  \centering
   \begin{tabular}{llccc}
           &            & negative & no effect & positive \\ \midrule
    {some} & relevance  & 0.019    & 0.592       & 0.389      \\
           & competence & 0        & 0         & 1        \\
           & prior      & 0.999        & 0.001     & 0    \\
    {or}   & relevance  & 0     & 0.228       & 0.772      \\
           & competence & 0      & 0.007       & 0.993      \\
           & prior      & 0.619    & 0.373       & 0.008      \\
   \end{tabular}
  \caption{Results of confirmatory analysis.}
  \label{tab:results-confirmatory}
\end{table}

Results from this analysis are summarized in Table \ref{tab:results-confirmatory}.
Consistent with predictions, for the trigger ``some'', we find a clear negative prior probabilty effect of the stronger alternative ``all'' being true, as inidcated by the probability of the negative effect of prior being $P = 0.999$ (Fig.\ref{posteriors-main}, \texttt{Prior~(some)} line, orange color). Similarly, we find a clear positive effect for speaker competence ($P =  1$, Fig.\ref{posteriors-main}, \texttt{Competence~(some)} line, green color). However, we do not find any clear effects of relevance (Fig.\ref{posteriors-main}, \texttt{Relevance~(some)} line, split colors). If anything, the data supports the hypothesis that the effect is not negative.

@@

\begin{figure}[h]
	\begin{center}
		\includegraphics[width=1\linewidth]{images/posterior-effects-categorical-noDupl.pdf}
	\end{center}
	\vspace{-0.3cm}
	\caption{Exploratory analysis with categorical predictors: Distributions of the posterior samples of each factor of interest for the trigger ``some'' and ``or''. The colors and the vertical dashed lines indicate the effect intervalls (from left to right: negative, no, positive effect). Points indicate posterior means, thick lines indicate 66\% credible intervals, thin lines indicate 95\% credible intervals. \mf{I'm not sure if we should blindly reuse a threshold of 0.05 here without further justification. Something to think about.}}
	\label{posteriors-cat}
\end{figure}

@@

\begin{figure}[h]
	\begin{center}
		\includegraphics[width=1\linewidth]{images/posterior-effects-means-noDupl.pdf}
	\end{center}
	\vspace{-0.3cm}
	\caption{Exploratory analysis with predictor ratings averaged by-vignette: Distributions of the posterior samples of each factor of interest for the trigger ``some'' and ``or''. The colors and the vertical dashed lines indicate the effect intervalls (from left to right: negative, no, positive effect). Points indicate posterior means, thick lines indicate 66\% credible intervals, thin lines indicate 95\% credible intervals. \mf{I'm not sure if we should blindly reuse a threshold of 0.05 here without further justification. Something to think about.}}
	\label{posteriors-mean}
\end{figure}

@@

\pt{We have a generally higher rate of implicature generations for ``some'' than for ``or'' (across conditions): the average inference likelihood rating for the former is 0.179, and -0.179 for the latter. Maybe something to mention in the discussion}
